---
title: "36-402 Homework 2"
author:
- Eu Jing Chua
- eujingc
date: "January 29, 2019"
output:
  pdf_document: default
header-includes:
    - \usepackage{enumerate}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

\newcommand{\est}[1]{\hat{#1}}
\newcommand{\betah}[1]{\est{\beta_{#1}}}
\newcommand{\avg}[1]{\overline{#1}}
\newcommand{\E}[1]{\mathbb{E} \left[ #1 \right]}
\newcommand{\Var}[1]{\text{Var} \left[ #1 \right]}
\newcommand{\Cov}[1]{\text{Cov} \left[ #1 \right]}
\newcommand{\X}{\mathbb{X}}
\newcommand{\sumTo}[1]{\sum^{#1}_{i=1}}
\newcommand{\sumjTo}[1]{\sum^{#1}_{j=1}}
\newcommand{\matr}[1]{\mathbf{#1}}

```{r}
library(knitr)
```

```{r}
data <- read.csv("uval.csv")
```

# Question 1

```{r}
lm.fit.1 <- lm(growth ~ underval + log(gdp), data = data)

kable(coef(summary(lm.fit.1)),
      digits = 5, caption = "Coefficients & Std. Error of linear model")
```

Since the coeficient of $log(gdp)$ is significantly positive, this model does not seem to support the idea of "catching-up" as countries with higher GDP have a higher economic growth rate. However, it does support the idea that under-valuing a currency boosts economic growth as the coefficient of $underval$ is quite significantly positive (assuming 5% significance level), indicating a positive $underval$ index, which represents undervaluing, leads to higher economic growth.

# Question 2

**Q2 a)**

```{r}
lm.fit.2 <- lm(growth ~ underval + log(gdp) + factor(country) + factor(year),
               data = data)

kable(coef(summary(lm.fit.2))[2:3, ],
      digits = 5, caption = "Coefficients & Std. Error of linear model")
```

**Q2 b)**
It is more appropriate to use `factor(year)` as there are only 10 unique years that are 5 years apart. As such, it might be more appropriate to model year as a discrete variable rather than a continuous one. Modelling this way, we will have a slope for each 5-year interval rather than a single slope for each increment of year.

**Q2 c)**

```{r}
plot(unique(data$year), c(0, coef(lm.fit.2)[182:190]),
     xlab = "Year", ylab = "Coefficient",
     main = "Plot of coefficient of each year across years")
```

**Q2 d)**

Since the coeficient of $log(gdp)$ is positive, this model does not seem to support the idea of "catching-up" as countries with higher GDP have a higher economic growth rate. However, it does support the idea that under-valuing a currency boosts economic growth as the coefficient of $underval$ is positive, indicating a positive $underval$ index, which represents undervaluing, leads to higher economic growth.

# Question 3

**Q3 a)**

```{r}
lm.r.squareds <- matrix(nrow = 2, ncol = 2)
rownames(lm.r.squareds) <- c("$R^2$", "Adj. $R^2$")
colnames(lm.r.squareds) <- c("Model 1", "Model 2")
lm.r.squareds[, 1] <- with(summary(lm.fit.1), c(r.squared, adj.r.squared))
lm.r.squareds[, 2] <- with(summary(lm.fit.2), c(r.squared, adj.r.squared))

kable(lm.r.squareds, digits = 5,
      caption = "$R^2$ values for each linear model")
```

**Q3 b)**

```{r}
# Taken from textbook chapter 3 page 77
cv.lm <- function(data, formulae, nfolds = 5) {
    data <- na.omit(data)
    formulae <- sapply(formulae, as.formula)
    n <- nrow(data)
    fold.labels <- sample(rep(1:nfolds, length.out = n))
    mses <- matrix(NA, nrow = nfolds, ncol = length(formulae))
    colnames <- as.character(formulae)
    for (fold in 1:nfolds) {
        test.rows <- which(fold.labels == fold)
        train <- data[-test.rows, ]
        test <- data[test.rows, ]
        for (form in 1:length(formulae)) {
            current.model <- lm(formula = formulae[[form]], data = train)
            predictions <- predict(current.model, newdata = test)
            test.responses <- eval(formulae[[form]][[2]], envir = test)
            test.errors <- test.responses - predictions
            mses[fold, form] <- mean(test.errors^2)
        }
    }
    return(colMeans(mses))
}
```

```{r cache=TRUE}
loocv.mse <- cv.lm(data, c("growth ~ underval + log(gdp)", "growth ~ underval + log(gdp) + factor(country) + factor(year)"), nfolds = nrow(data))
```

```{r}
names(loocv.mse) <- c("Model 1", "Model 2")
kable(loocv.mse, caption = "$\\hat{MSE}$ of linear models by LOOCV")
```

The second linear model seems to perform better on predictions, having an estimated MSE that is `r signif((loocv.mse[1] - loocv.mse[2]) / loocv.mse[1] * 100, 3)`% lower than the first model.

**Q3 c)**

Since the lowest count of any country is 2, doing 5-fold cross validation has a high chance of resulting in the training 4-folds missing some countries because the 2 rows are in the testing 1-fold that is left out. This would create models that leave out some countries by chance and fail to predict with a new category it has not seen in training.

# Question 4

```{r}
library(np)
options(np.messages = FALSE)
```
**Q4 a)**

```{r cache=TRUE}
npreg.fit <- npreg(growth ~ log(gdp) + underval + year, data = data,
                   tol = 10^-3, ftol = 10^-4)

```

There are no coefficients to report for kernel regression, as it is a non-parametric smoothing method that smooths the data with a kernel. The smoothing is only controlled by the choice of the kernel and the related bandwidth for the kernel, without any coefficients on the predictors.

**Q4 b)**

```{r}
plot(fitted(npreg.fit) ~ data$growth,
     xlab = "Observed Growth", ylab = "Predicted Growth",
     main = "Plot of Predicted Growth against Observed Growth")
```

**Q4 c)**

```{r}
plot(resid(npreg.fit) ~ fitted(npreg.fit),
     xlab = "Predicted Growth", ylab = "Residuals",
     main = "Plot of Residuals against Predicted Growth")
abline(h = 0)
```

The points should be scattered around a flat line at 0 if the model was right, as we assume $\E{\epsilon} = 0$.
In this case, it would seem that the residuals indeed are roughly scattered around a flat line at 0.

**Q4 d)**

```{r cache=TRUE}
lm.mse <- cv.lm(data, c("growth ~ underval + log(gdp) + factor(year)"), nfolds = nrow(data))
```
```{r}
results <- c(lm.mse, npreg.fit$bws$fval)
names(results) <- c("Linear Regression", "Kernel Regression")
kable(results, digits = 7, caption = "$\\hat{MSE}$ of the regressions via cross validation")
```

As seen from above, kernel regression has a lower estimated MSE than a linear model with the same covariates. Hence, the kernel regression is better in generalizing and predicting better than the linear model.

# Question 5

**Q5 a)**

```{r fig.height = 4}
years <- seq(from = 1955, to = 2000, by = 5)
n <- length(years)
plot(predict(npreg.fit, newdata = data.frame(underval = rep(0, n), gdp = rep(20000, n), year = years)) ~ years,
     xlab = "Year", ylab = "Growth",
     main = "Predicted growth with gdp = 20000, underval = 0")
```

**Q5 b)**

```{r fig.height = 4}
plot(predict(npreg.fit, newdata = data.frame(underval = rep(0.5, n), gdp = rep(20000, n), year = years)) ~ years,
     xlab = "Year", ylab = "Growth",
     main = "Predicted growth with gdp = 20000, underval = +0.5")
```

**Q5 c)**

```{r fig.height = 4}
plot(predict(npreg.fit, newdata = data.frame(underval = rep(0, n), gdp = rep(3000, n), year = years)) ~ years,
     xlab = "Year", ylab = "Growth",
     main = "Predicted growth with gdp = 3000, underval = 0")
```

**Q5 d)**

```{r fig.height = 4}
plot(predict(npreg.fit, newdata = data.frame(underval = rep(0.5, n), gdp = rep(3000, n), year = years)) ~ years,
     xlab = "Year", ylab = "Growth",
     main = "Predicted growth with gdp = 3000, underval = +0.5")
```

**Q5 e)**

By comparing the graphs with initial $GDP = 20000$, we can see that by increasing under-valuation from 0 to 0.5, the predicted values increased from 1955 to 1970 but for $GDP = 3000$, increasing under-valuation from 0 to 0.5 decreased the predicted values from 1955 to 1970. The graphs are not parallel across and the effects not similar when we change intial GDP and under-valuation, hence there should be some interaction between these two variables.

**Q5 f)**

```{r fig.height = 10, fig.width = 8}
plot(npreg.fit)
```

From the plots of growth against each variable, we can see that both GDP and under-valuation have strong relationshipswith growth. For the plot of growth against $log(GDP)$, there seems to be significant variation from varying $log(GDP)$ from 7 to 11. For the plot of growth against under-valuation, there also seems to be significant variation from varying under-valuation from -2.0 to -0.5. Finally, there is also some relationship, a weaker one than the previous two, between growth and year. For the plot of growth against year, there seems to be a slightly significant variation from avarying year from 1970 to 1985.

# Question 6

\begin{align}
\text{Optimism} &= \E{\frac{1}{n} \sumTo{n} (Y_i' - \est{\mu}(x_i))^2} - \E{\frac{1}{n} \sumTo{n} (Y_i - \est{\mu}(x_i))^2} \\
    &= \frac{1}{n} \sumTo{n} \E{(Y_i' - \est{\mu(x_i)})^2} - \E{(Y_i - \est{\mu(x_i)})^2} \\
    &= \frac{1}{n} \sumTo{n} \Var{Y_i' - \est{\mu}(x_i)} + \E{Y_i' - \est{\mu(x_i)}}^2 - \Var{Y_i - \est{\mu(x_i)}} - \E{Y_i - \est{\mu(x_i)}}^2 \\
    &= \frac{1}{n} \sumTo{n} \Var{Y_i'} + \Var{\est{\mu}(x_i)} - 2 \Cov{Y_i', \est{\mu}(x_i)} - \Var{Y_i} - \Var{\est{\mu}(x_i)} + 2 \Cov{Y_i, \est{\mu}(x_i)} \\
        &+ (\E{Y_i'} - \E{\est{\mu}(x_i)})^2 - (\E{Y_i} - \E{\est{\mu}(x_i)})^2 \\
    &= \frac{1}{n} \sumTo{n} 2 \Cov{Y_i, \est{\mu}(x_i)}, \hspace{0.05in} \text{as $\Var{Y_i'} = \Var{Y_i}$, $\E{Y_i'} = \E{Y_i}$, and $\Cov{Y_i', \est{\mu}(x_i)} = 0$} \\
    &= \frac{2}{n} \sumTo{n} \Cov{Y_i, \sumjTo{n} w(x_j, x_i) Y_j} \\
    &= \frac{2}{n} \sumTo{n} w(x_i, x_i) \Var{Y_i}, \hspace{0.05in} \text{as} \hspace{0.05in} \Cov{Y_i, Y_j} = 0 \forall i \ne j \\
    &= \frac{2 \sigma^2}{n} \text{tr}(\matr{w}) = \frac{2 \sigma^2}{n} df(\est{\mu})
\end{align}

# Question 7

\begin{align}
\est{h_{CV}} &= h_{opt} O(n^{-1/10}) + 2 h_{opt} \\
    &= O(n^{-1/5}) O(n^{-1/10}) + O(n^{-1/5}) \\
    &= O(n^{-3/10}) + O(n^{-1/5}) \\
    &= O(n^{-1/5}) \\
MSE(\est{h_{CV}}) - \sigma^2(x) &= O(\est{h_{CV}}^4) + O((n \est{h_{CV}})^{-1}) \\
    &= O((n^{-1/5})^4) + O(((n^{1 - 1/5}))^{-1}) \\
    &= O(n^{-4/5}) + O(n^{-4/5}) \\
    &= O(n^{-4/5})
\end{align}
